---
title: "Exploratory Data Analysis on Game Sales in R"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, comment = NA, warning = F)
```

## Exploratory Data Analysis on Game Sales in R

The data analysis pipeline begins with importing or creating an analysis-ready dataset. 
 
```{r}
#load packages to be used

pacman::p_load(readxl, tidyverse, reshape2, knitr, dplyr, ggplot2, ftExtra,scales,knitr,ggthemes,plotly, kableExtra, colorDF, lubridate, janitor, ggrepel)

#load colours to be used in vizualizations 

colours <- c("#D58A94","#16E2F5", "#560319", "#52595D","#151B54","#006A4E","#E2F516", "#DC143C", "#01F9C6", "#9C7F61", "#50C878", "#513B1C","#00BFFF", "#7D0552","#008000", "#9DD0C0", "#BE9D7C", "#A2C8B9","#583759", "#F6BE00", "#F5F5F5", "#8B008B", "#5865F2", "#FA2A55", "#F88017", "#ADF802", "#0002FF") 

```

```{r}

#import clean dataset

data <- read.csv("clean game sales.csv")

#get idea of data as a neat table

head(data) %>% kbl() %>% kable_material(c("striped", "hover"))
```

### Stage 1 - Summarizing

This stages will help us understand the data through descriptive statistics. The benefits of this stage are as follows:

 * we will have a basic information about variables in our dataset.
 * we will get an idea of potential relationships between variables.
 
```{r}
#summarize each varibable in the dataset

summary(data) %>% kbl %>% kable_material(c("striped", "hover"))

```

```{r}
#categorical



```


### TRansforming The Data

```{r}

print(unique(data$Platform))

```

```{r}

data_trans  <- data %>% mutate(game = case_when(Platform == "PS3" |Platform == "PS4" | Platform == "PS2" | Platform == "PSP" ~ "Playstation",
                                                
                                 TRUE ~ "Other"))

head(data_trans)                  
```

### Stage 3 - Vizualizing The Data


```{r}
#create a histogram of the variables

 mutate(age_groups = ifelse(age<25, "18 - 24",
                             ifelse(age<35, "25 - 34",
                                    ifelse(age<45, "35 - 45", "45+"))))

data %>%
  mutate(grouped_platfroms = ifelse(Platform =  ) )
  group_by(Platform) %>%
  tally() %>%
  adorn_percentages('col') %>%
  mutate(perc = paste(round(n*100, 2), "%")) %>%
  ggplot(aes(x = "", y = n, fill = Platform, label = perc)) +
  geom_bar(stat = "identity", position = "fill") +
  geom_text(position = position_fill(vjust = 0.5), size = 2.5) +
  coord_polar("y") +
  theme_void() +
  labs(title = "Distribution Of Platform") #+ scale_fill_manual(values = c(colours))


#frequency on Platform
ggplot(data, aes(x = Platform)) +
  geom_bar(bins = 100, fill = colours[5]) +
  labs(title = "Frequency of Sales By Platfrom ", x = "Platform", y ="count") +
  theme_economist()
  

```
#### Year

```{r}
#pie chart



#Frequency on years

ggplot(data, aes( x = Year)) +
  geom_histogram(bins = 100, fill = colours[5]) +
  labs(title = "Histgram Of Frequency of Sales Yearly ", x = "Year", y ="count") +
  theme_economist()

```

#### Genre

```{r}

#pie chart

data %>%
  group_by(Genre) %>%
  tally() %>%
  adorn_percentages('col') %>%
  mutate(perc = paste(round(n*100, 2), "%")) %>%
  ggplot(aes(x = "", y = n, fill = Genre, label = perc)) +
  geom_bar(stat = "identity", position = "fill") +
  geom_text(position = position_fill(vjust = 0.5), size = 2.5) +
  coord_polar("y") +
  theme_void() +
  labs(title = "Distribution Of Genre") + 
  scale_fill_manual(values = c(colours))

#frequency on Genre
ggplot(data, aes(x = Genre)) +
  geom_bar(bins = 100, fill = colours[5]) +
  labs(title = "Frequency of Sales By Genre ", x = "Genre", y ="count") +
  theme_economist()
  
```



